---
title: "Final Library Design"
author: Shannon Laub
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if(!require('pacman')) {install.packages('pacman')}
pacman::p_load(openxlsx, dplyr, tidyr, reshape2, stringr)

# BiocManager::install("GenomicRanges")
library(GenomicRanges)

source('libraryDesignFunctions.R')
```

Before running this script, go through the following to generate guide pools:
- preprocessCRISPick.Rmd - to create raw guide pools
- preprocessGDO.Rmd - to preprocess the raw guide pools for this script

# Set Directory Information
```{r}
wd <- getwd()
wd # verify directory location

# Set universal directories:
dirRegion <- paste0(wd,"//preprocessed//")
dirGDO <- paste0(wd, "//preprocessed//")

snpList <- unique(read.xlsx(paste0(wd,"//input//snp_loc_hg19.xlsx")))
rownames(snpList) <- NULL
snpList$chr <- as.character(snpList$chr)
```

# Assemble Library
```{r}
regionList <- c("regions200", "regions300", "regions400", "regions500", "regions600")
fileList <- c("200bp_i30","300bp_i30","400bp_i30","500bp_i30","600bp_i30")
targetGDO <- assembleLibrary(3, fileList, regionList, dirGDO, dirRegion,
                          onTargetFilter = 0.2, pickByStrand = TRUE, overlapLeniency=2)

uniqueGDO <- unique(targetGDO)
nrow(uniqueGDO) == nrow(targetGDO)

overlap <- findSelfOverlap(targetGDO, "chr", "start","end", selfBool=TRUE)
overlapLenient <- findSelfOverlap(targetGDO, "chr", "startL","endL", selfBool=TRUE)
```

```{r}
qcSimple(targetGDO)
qcSumOff(targetGDO)
orient <- qcStrand(targetGDO)
```


Quality Control
```{r}
# Load regions for reference

regions200 <- read.xlsx(paste0(wd,"//preprocessed//regions200.xlsx"))
regions500 <- read.xlsx(paste0(wd,"//preprocessed//regions500.xlsx"))

# Check that you have all the expected regions from the 200bp
  # i.e. you haven't LOST any regions and you haven't ADDED any regions by accident

all(targetGDO$SNP %in% regions200$SNP)
!any(is.na(targetGDO$SNP))

# Check that each guide overlaps with only ONE REGION

reg200Overlap <- findOverlap(targetGDO, regions200,
                          "chr", "start", "end",
                          "chr", "start", "end")
# 201 out of 222 guides overlap the 200bp region

reg500Overlap <- findOverlap(targetGDO, regions500,
                          "chr", "start", "end",
                          "chr", "start", "end")
# All guides overlap a 500bp region.
```

Export guide list:
```{r}
write.xlsx(targetGDO, file = paste0(wd, "//output//finalTargetGDOi30v2.xlsx"))

# targetGDO <- read.xlsx(paste0(wd,"//output//finalTargetGDOi30.xlsx"))
```

# Updating Positive Controls
We submitted a list of interesting genes to CRISPick to design promoter targeting control guides.
Next, we want to:
1) Remove any designed guides that overlap with our validated guides.
2) Of the remaining pool, pick the best performing guides. (i.e. top 3 or top 5)

## Load Data
```{r}
wd <- getwd()

# Import new guide designs
posA30 <- read.xlsx(paste0(wd,"//crispick_output//sgrna_designs_promoter_a30.xlsx"))
names(posA30)[names(posA30)=="Target.Gene.Symbol"] <- "SNP"

# Get refseq conversion for chromosome numbers. Needed for compatibility with CRISPick
refseqLookup <- read.delim(paste0(wd,"//input//seq_report_hg19.tsv"))
refseqLookup <- subset(refseqLookup, Role=="assembled-molecule")

# Rename columns for convenience
names(refseqLookup)[names(refseqLookup)=="Chromosome.name"] <- "chr"
names(refseqLookup)[names(refseqLookup)=="RefSeq.seq.accession"] <- "refseq"
refseqLookup <- refseqLookup %>%
  select(chr, refseq)
```

```{r}
# Add chromosome number based on refseq lookup table
posA30 <- addGDOLoc(posA30, refseqLookup)

# Use Input as the "SNP" region
names(posA30)[names(posA30)=="Target.Gene.Symbol"] <- "SNP"

all(posA30$start < posA30$end) # CHECK that coordinate order is correct. TRUE if okay.

# Add rownumber for reference later
posA30$n <- as.numeric(rownames(posA30))


all(posA30$n == as.numeric(rownames(posA30))) #CHECK that n == rownumber. TRUE if okay.
```

```{r}
posGene <- read.table(paste0(wd,"//input//crispick_pos_ctrl.txt"))
colnames(posGene) <- "SNP"

validCtrl <- read.xlsx(paste0(wd, "//input//validated_pos_ctrl.xlsx"))
```

First remove any guides that overlap the validated guides.

```{r}
posOverlap <- findOverlap(validCtrl, posA30,
                          "chr", "start", "end",
                          "chr", "start", "end")
# 12 new guides overlap the validated guides

# Remove guides that overlap validated guides from the pool.
posPool <- posA30[!posA30$n %in% posOverlap[,2],]

```

Pick top 3 guides for each promoter region:
```{r}
# Get indices of guides that are overlapping.
posOverlap <- findSelfOverlap(posPool, "chr", "start", "end")

# Pick top guides for each promoter region.
topPos <- pickTopGDO(3, posA30, posGene, posOverlap)
```

Export the top guides. Go in order of pick order, fill in where we don't have validated guides.
Will combine the lists manually since there's not too many guides.
Need to manually pick+evaluate RPA1/SMYD4 guides since the region is overlapping..
```{r}
write.xlsx(topPos, file = paste0(wd, "//output//top3_promoter_a30.xlsx"))
```

